Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-10 paper.qmd
Updated Figure 3

nldr_two_curvy <- tsne_two_curvy_scaled |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)

@@

hex_grid_two_curvy <- ggplot(
  data = hex_grid_with_counts, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = tsne_two_curvy_scaled, 
             aes(x = tSNE1, y = tSNE2), 
             alpha = 0.5, size = 0.5, color = "#636363") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("b", sc_ltr_pos)

@@

hex_grid_nonempty_two_curvy <- ggplot(
  data = hex_grid_nonempty, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = df_bin_centroids_two_curvy, 
             aes(x = c_x, y = c_y), 
             color = "#000000",
             size = 1) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos) 
```

@@

trimesh_removed_two_curvy <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  coord_equal() +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("d", sc_ltr_pos)

@@

trimesh_removed_mnist <- ggplot() + 
  geom_segment(data = tr_from_to_df_mnist, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_minst_scaled_temp,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c("#999999",'#ff7f00')) +
  interior_annotation("a", sc_ltr_pos_mnist) +
  theme(legend.position = "none")

tsne_plot_mnist_clust <- ggplot() + 
  geom_point(data = tsne_minst_scaled_temp,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
  interior_annotation("a", sc_ltr_pos_mnist) +
  theme(legend.position = "none")

@@

mnist_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5,
    alpha = 0.5) +
  geom_point(
    size = 0.5,
    alpha = 0.5) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1.1, 0.9)) +
  ylim(c(-1.1, 0.9)) +
  interior_annotation("b") +
  theme(
    legend.position = "none"
  )

mnist_proj_tsne_model1_link <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5,
    alpha = 0.5) +
  geom_point(
    size = 1.2,
    alpha = 0.5) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1.1, 0.9)) +
  ylim(c(-1.1, 0.9)) +
  interior_annotation("b") +
  theme(
    legend.position = "none"
  )

@@

mnist_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5) +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c") +
  theme(
    legend.position = "none"
  )

mnist_proj_tsne_model2_link <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5) +
  geom_point(
    size = 1.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c") +
  theme(
    legend.position = "none"
  )

@@

mnist_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    #size = 0.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 0.6)) +
  ylim(c(-1, 0.6)) +
  interior_annotation("d") +
  theme(
    legend.position = "none"
  )

>>>>

nldr_two_curvy <- tsne_two_curvy_scaled |> 
  ggplot(aes(x = tSNE1, y = tSNE2)) + 
  geom_point(alpha=0.5, colour="#636363", size = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos, cex = 2)

@@

hex_grid_two_curvy <- ggplot(
  data = hex_grid_with_counts, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = tsne_two_curvy_scaled, 
             aes(x = tSNE1, y = tSNE2), 
             alpha = 0.5, size = 0.5, color = "#636363") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("b", sc_ltr_pos, cex = 2)

@@

hex_grid_nonempty_two_curvy <- ggplot(
  data = hex_grid_nonempty, 
  aes(x = x, y = y)) +
  geom_polygon(color = "grey70", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = df_bin_centroids_two_curvy, 
             aes(x = c_x, y = c_y), 
             color = "#000000",
             size = 1) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos, cex = 2) 
```

@@

trimesh_removed_two_curvy <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  coord_equal() +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("d", sc_ltr_pos, cex = 2)

@@

trimesh_removed_mnist <- ggplot() + 
  geom_segment(data = tr_from_to_df_mnist, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = tsne_minst_scaled_temp,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c("#999999",'#ff7f00')) +
  interior_annotation("a", sc_ltr_pos_mnist, cex = 2) +
  theme(legend.position = "none")

tsne_plot_mnist_clust <- ggplot() + 
  geom_point(data = tsne_minst_scaled_temp,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
  interior_annotation("a", sc_ltr_pos_mnist, cex = 2) +
  theme(legend.position = "none")

@@

mnist_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5,
    alpha = 0.5) +
  geom_point(
    size = 0.5,
    alpha = 0.5) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1.1, 0.9)) +
  ylim(c(-1.1, 0.9)) +
  interior_annotation("b", cex = 2) +
  theme(
    legend.position = "none"
  )

mnist_proj_tsne_model1_link <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5,
    alpha = 0.5) +
  geom_point(
    size = 1.2,
    alpha = 0.5) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  #coord_fixed() +
  xlim(c(-1.1, 0.9)) +
  ylim(c(-1.1, 0.9)) +
  interior_annotation("b", cex = 2) +
  theme(
    legend.position = "none"
  )

@@

mnist_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5) +
  geom_point(
    size = 0.5,
    alpha = 0.3) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c", cex = 2) +
  theme(
    legend.position = "none"
  )

mnist_proj_tsne_model2_link <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    linewidth = 0.5) +
  geom_point(
    size = 1.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 1)) +
  ylim(c(-1, 1)) +
  interior_annotation("c", cex = 2) +
  theme(
    legend.position = "none"
  )

@@

mnist_proj_tsne_model3 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2,
      colour = cluster)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    #size = 0.2,
    alpha = 0.3) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 2) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1, 0.6)) +
  ylim(c(-1, 0.6)) +
  interior_annotation("d", cex = 2) +
  theme(
    legend.position = "none"
  )
